---
title: "Project Assignment 3"
author: "David Abeiku Saah"
date: "02/04/2023"
output: word_document
---

```{r setup, message=FALSE, warning=FALSE}
```

### Research Question

**Are there significant differences in levels of trust in government and private media sources?**

Explanatory variables: Public media, government, private media.
Response variable: Level of trust.

All the variables are ***categorical***.

## 1. Load data set(s) and libraries

```{r}
# load dataset and library packages
load("Ghana_r8");
library(descr);
library(stats);
``` 


## 2. Create variable subset

```{r}
# assign the variables to be used
vars = c("Q41N_GHA", "Q41O_GHA", "Q41M_GHA");

# create a new data frame with the variables
my_data = data.frame(Ghana_r8[vars]);
```

## 3. Data management I: check for and recode errors and NAs

```{r}
# Check the graph if there are any error codes
freq(my_data$Q41N_GHA, cex.names=0.55);

# "Missing", "Refused" and "Don't know" are NA values since they identify as missing data
my_data$Q41N_GHA[my_data$Q41N_GHA == "Refused"] = NA;
my_data$Q41N_GHA[my_data$Q41N_GHA == "Missing"] = NA;
my_data$Q41N_GHA[my_data$Q41N_GHA == "Don’t know"] = NA;
my_data$Q41N_GHA = factor(my_data$Q41N_GHA);
freq(my_data$Q41N_GHA, cex.names=0.8);

# Check the graph if there are any error codes
freq(my_data$Q41O_GHA, cex.names=0.55);

# "Missing", "Refused" and "Don't know" are NA values since they identify as missing data
my_data$Q41O_GHA[my_data$Q41O_GHA == "Refused"] = NA;
my_data$Q41O_GHA[my_data$Q41O_GHA == "Missing"] = NA;
my_data$Q41O_GHA[my_data$Q41O_GHA == "Don’t know"] = NA;
my_data$Q41O_GHA = factor(my_data$Q41O_GHA);
freq(my_data$Q41O_GHA, cex.names=0.8);

# Check the graph if there are any error codes
freq(my_data$Q41M_GHA, cex.names=0.55);

# "Missing", "Refused" and "Don't know" are NA values since they identify as missing data
my_data$Q41M_GHA[my_data$Q41M_GHA == "Refused"] = NA;
my_data$Q41M_GHA[my_data$Q41M_GHA == "Missing"] = NA;
my_data$Q41M_GHA[my_data$Q41M_GHA == "Don’t know"] = NA;
my_data$Q41M_GHA = factor(my_data$Q41M_GHA);
freq(my_data$Q41M_GHA, cex.names=0.8);
```

## 4. Data management II: further subset and create secondary variable

```{r}
# Merge "Just a little" and "Somewhat" into "Somewhat trust"

# create new variables
my_data$public_media <- rep(NA,nrow(my_data));
my_data$private_media <- rep(NA,nrow(my_data));
my_data$government <- rep(NA,nrow(my_data));

# public media
my_data$public_media[my_data$Q41N_GHA == "Not at all"] <- "Not at all";
my_data$public_media[my_data$Q41N_GHA == "Just a little"] <- "Somewhat trust";
my_data$public_media[my_data$Q41N_GHA == "Somewhat"] <- "Somewhat trust";
my_data$public_media[my_data$Q41N_GHA == "A lot"] <- "A lot";
freq(my_data$public_media);

# private media
my_data$private_media[my_data$Q41O_GHA == "Not at all"] <- "Not at all";
my_data$private_media[my_data$Q41O_GHA == "Just a little"] <- "Somewhat trust";
my_data$private_media[my_data$Q41O_GHA == "Somewhat"] <- "Somewhat trust";
my_data$private_media[my_data$Q41O_GHA == "A lot"] <- "A lot";
freq(my_data$private_media);

# government sources

my_data$government[my_data$Q41M_GHA == "Not at all"] <- "Not at all";
my_data$government[my_data$Q41M_GHA == "Just a little"] <- "Somewhat trust";
my_data$government[my_data$Q41M_GHA == "Somewhat"] <- "Somewhat trust";
my_data$government[my_data$Q41M_GHA == "A lot"] <- "A lot";
freq(my_data$government);
```

## 5. Descriptive statistics (sample means, standard deviations, proportions) and univariate displays

```{r}
# Univariate display for public media information source
freq(my_data$public_media, main = "Ghanaians' Trust in Public News Media", y.axis = "percent", ylab = "Percentage of People", xlab = "Level of Trust");

# Univariate display for private media information source
freq(my_data$private_media, main = "Ghanaians' Trust in Private News Media", y.axis = "percent", ylab = "Percentage of People", xlab = "Level of Trust");

# Univariate display for government information source
freq(my_data$government, main = "Ghanaians' Trust in Government Information Sources", y.axis = "percent", ylab = "Percentage of People", xlab = "Level of Trust");
```

## 6. Bivariate tables and graphs

```{r}

```

I have learnt that most Ghanaians are indecisive of what type of information they believe or trust. The proportion of Ghanaians from the sample that are indecisive are 69.48%, 69.93% and 67.36% for public media, private media and government information sources. A few percentage of people made a choice to accept or reject information as truth depending on its source. 

It is possible that there is no significant difference between the level of trust in government and private media sources.

## 7. Bivariate analysis (hypothesis tests and post-hoc tests)

```{r}
```

## 8. Moderation

```{r}
```

## 9. Save

```{r}
# save the data
save(my_data, file = "./my_data.RData");
```